---
title: "Caring Analysis"
author: "Jonathan Luu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = F)
library(tidyverse)
library(readxl)
library(lubridate)

# Read in data
path <- "/Users/jonathanluu/Library/CloudStorage/Dropbox/Caring Health/Data/Caring_Data_V2.xlsx"
dat <- lapply(excel_sheets(path = path), function(x) 
  read_excel(path = path, sheet = x, col_types = c(rep("guess",20),"text","logical")))
dat <- bind_rows(dat)

# Create categorical visit
dat <- dat %>% mutate(cat_visit = case_when(num_of_visits <= 1 ~ 0, 
                                 num_of_visits > 1 & num_of_visits <= 3 ~ 1,
                                 num_of_visits > 3 & num_of_visits <= 10 ~ 2,
                                 num_of_visits > 10 & num_of_visits <= 25 ~ 3,
                                 num_of_visits > 25 & num_of_visits <= 50 ~ 4,
                                 num_of_visits > 50 ~ 5))

# Create categorical age
dat <- dat %>% mutate(cat_age = case_when(Age <= 39 ~ 0, 
                                 Age > 39 & Age <= 64 ~ 1,
                                 Age > 65 ~ 2))

# Rename some columns
dat <- rename(dat, Language = "LANGUAGE_NAME", Ethnicity = "ETHNIC_GROUP", Race = "RACE")

# Format dates
dat <- dat %>% mutate(ENC_DATE = ymd(ENC_DATE), Last_Visit_Date = ymd(Last_Visit_Date))

# Get distinct IDs
dat.distinct <- dat %>% distinct(PAT_ID, .keep_all = T)
```

# Exploratory analysis 

This dataset is based off version 1 of the dataset. It contains encounter visits between 05/2018 to 10/2022 using the following inclusion/exclusion criteria:

* Visit date during the reporting period (determined by report parameters)
* Visit took place in an SA38 department
* Visit has at least one non-voided charge
* Patient is 18 years or older as of the date the report is run

For the first analysis, I explored the dataset to get a better understanding of the number of observations, the distributions of all the variables, and extent of missingness. 

\newpage
### Age

```{r}
# Number of unique entries in the data
length(unique(dat.distinct$PAT_ID))

# Age distribution
dat.distinct$Age %>% summary()
```

```{r, echo=FALSE}
dat.distinct$Age %>% hist(main="Histogram of age", xlab="Age (years)")
```

\newpage
### Language

```{r}
# Language
dat.distinct$Language %>% unique() %>% length() # 67 different languages spoken
dat.distinct$Language %>% table() %>% sort(decreasing = T) %>% 
  head(n=15) %>% as_tibble() %>% rename(Language=".") # Most popular languages
```

\newpage
### Race

```{r, echo=FALSE}
# Race
race.table <- dat.distinct$Race %>% table() %>% sort() %>% 
  as.data.frame() %>% rename(Race=".")
race.table %>% as_tibble()

# Race barplot
race.table %>% ggplot(aes(x=Freq, y=Race)) + geom_bar(stat="identity") +
  theme(axis.text.y=element_text(angle=0,hjust=1,vjust=0.5)) + 
  ggtitle("Race frequency")
```

Again, we can see there are some missing data for race.

\newpage
### Ethnicity

```{r, echo=FALSE}
# Ethnicity
eth.table <- dat.distinct$Ethnicity %>% table() %>% sort() %>% 
  as.data.frame() %>% rename(Ethnicity=".")
eth.table %>% as_tibble()

# Ethnicity barplot
eth.table %>% ggplot(aes(x=Freq, y=Ethnicity)) + geom_bar(stat="identity") +
  theme(axis.text.y=element_text(angle=0,hjust=1,vjust=0.5)) + 
  ggtitle("Ethnicity frequency")
```

Similarly for ethnicity, we have around 3130 unknown values and 90 refusals. 

\newpage
### Number of visits

```{r}
# Number of visits
dat.distinct$num_of_visits %>% summary()

# Most common number of visits
dat.distinct$num_of_visits %>% table() %>% sort(decreasing=T) %>% 
  head(n=10) %>% as_tibble() %>% rename("Num Visit"=".")
```

```{r,echo=FALSE}
dat.distinct %>% filter(num_of_visits <= 15) %>% pull(num_of_visits) %>% 
  hist(breaks=0:15, main="Histogram of number of visits", xlab="Number of visits")
```

\newpage
### PCP information

We can see the distribution of PCPs seeing each of these individuals, as well as the distribution of PCP degrees

```{r, echo=FALSE}
dat$PCP_NAME %>% table() %>% sort(decreasing=T) %>% 
  as.tibble() %>% rename("PCP Name"=".") %>% print(n=25)

dat$PCP_DEGREE %>% table() %>% sort(decreasing=T) %>% 
  as.tibble() %>% rename("PCP Degree"=".")
```

\newpage
### Provider types and charged visits

Lastly, the distributions of provider types that took in and entered in data, as well as whether the visit is considered a charged-visit.

```{r, echo=FALSE}
dat$TakenProviderType %>% table() %>% sort(decreasing=T) %>% 
  as.tibble() %>% rename("Taken Provider Type"=".") # Provider type distribution
dat$HasChargeVisit %>% table() %>% sort(decreasing=T) %>% 
  as.tibble() %>% rename("Charged visit"=".") # Charged visits
```


\newpage
# Survey question descriptives 

Next, I looked at the distribution of response types - what were the most common questions being answered, what categories did these responses fall into, and which individuals had the most number of responses.

```{r}
# Number of response questions being asked
dat$MEAS_DISP_NAME %>% unique() %>% length() # 45 different response questions
options(pillar.print_max = 50, pillar.print_min = 50)
dat$MEAS_DISP_NAME %>% table() %>% sort(decreasing=T) %>% 
  as_tibble() %>% rename(Question=".")

# Domain categories
dat$DOMAIN_GRP %>% table() %>% sort(decreasing=T) %>% 
  as_tibble() %>% rename(Response=".") # 13 categories of responses
```

Note that we have an empty domain, indicating the individuals who did not answer a particular question.

```{r}
# Individuals with the most # of responses
dat$PAT_ID %>% table() %>% sort(decreasing = T) %>% head(n=10) %>% 
  as_tibble() %>% rename(ID=".")
```

From this, we can see some individuals have a very high number of responses.

\newpage
# Screening and missingness analysis

Next, I looked at how many individuals answered at least one SDH screening. Among those, what were the distributions that answered at least one financial question, one relationship safety question, and all domain questions. These numbers are summarized below:

```{r}
# 32421 unique individuals 
# 335402 total # of visits ranging from 1 to 379 visits per individual
# 266077 observations - mix of single observation without SDH and 1+ observations with SDH
# 242597 screening questions answered
# 23480 individuals with no SDH information

# 8941 individuals answered at least one question
# 12964 unique encounters 
## 6012 individuals answered a financial question (67% of 8941, 19% of 32421)
## 8135 individuals answered a relationship safety question (91% of 8941, 25% of 32421)
## 5996 individuals with 4 domain questions answered (67% of 8941, 19% of 32421)
  ## Housing, transportation, food, and stress
## 2120 individuals with all domain questions answered (24% of 8941, 7% of 32421)
```

From these numbers, we can see there is a mix of individuals for whom there are no problems and those who have SDH challenges but do not respond.

Note: 15 individuals with 0 visits

```{r, echo=FALSE}
dat.distinct %>% filter(num_of_visits==0) %>% 
  select(PAT_ID, num_of_visits, Last_Visit_Date, ENC_DATE) %>% print(n=15)
```


```{r, include=FALSE}
# Create new dataset showing number of responses by id and encounter
dat.resp.enc <- dat %>% count(PAT_ID, DOMAIN_GRP, ENC_DATE) %>% mutate_all(na_if,"") %>%
  pivot_wider(names_from = DOMAIN_GRP, values_from = n) 

# Ranges of screening
dat.resp.enc <- dat.resp.enc %>% mutate_if(is.integer, ~replace(.,is.na(.),0))
dat.resp.enc <- dat.resp.enc %>% select(-"NA") %>% mutate(sum = rowSums(across(where(is.numeric))))

# Number of questions answered
dat.resp.enc %>% filter(sum==0) %>% select(PAT_ID) %>% unique() %>% nrow() # 23480 people did not have a survey answer
ans.n <- dat.resp.enc %>% filter(sum > 0) %>% select(PAT_ID) %>% unique() %>% nrow() # 8941 individuals who did answer

# Extract covariates for each ID
covar <- dat.distinct %>% select(PAT_ID, cat_age, cat_visit, Race, Ethnicity, Language, HasChargeVisit)
dat.resp.enc <- merge(dat.resp.enc,covar,all.x=T)

# What constitutes completion? - Distribution of those who did complete a survey and how many questions did they answer?
# 6012 people answered a financial question
total.n <- dat.distinct %>% nrow()
fin.n <- nrow(dat.resp.enc %>% filter(`Financial/Other`>0) %>% select(PAT_ID) %>% unique())
fin.n/ans.n # 67% of the 8941 who answered
fin.n/total.n # 18% of the total 32421

# 8135 people answered a relationship safety question
rel.n <- nrow(dat.resp.enc %>% filter(`Relationship Safety`>0) %>% select(PAT_ID) %>% unique())
rel.n/ans.n # 91% of the 8941 who answered
rel.n/total.n # 25% of the total 32421

# People who answered all questions in these 4 categories: Housing, transportation, food, and stress
group.n <- dat.resp.enc %>% filter(sum>0)
group.n.id <- dat.resp.enc %>% select("Food Insecurity","Housing","Transportation","Stress and Social Isolation") > 0
group.n.id <- which(tibble(group.n.id) %>% apply(1,sum) == 4)
group.n <- dat.resp.enc %>% slice(group.n.id)
nrow(group.n) # 7774 encounters with these 4 domain questions answered
length(unique(group.n$PAT_ID)) # 5996 individuals with these 4 domain questions answered

# People who answered a question in all domains
dom.n <- dat.resp.enc %>% filter(sum>0)
dom.n.id <- dom.n %>% select(-c(PAT_ID,ENC_DATE,sum)) > 0
dom.n.id <- which(tibble(dom.n.id) %>% apply(1,sum) == 12)
dom.n <- dom.n %>% slice(dom.n.id)
nrow(dom.n) # 2332 encounters with all domain questions answered
length(unique(dom.n$PAT_ID)) # 2120 individuals with all domain questions answered

# How many screens are done per individual
# Assumption: A screen is when at least 1 question is answered during an encounter
screen.count <- dat.resp.enc %>% filter(sum>0) %>% count(PAT_ID)
summary(screen.count$n)
table(screen.count$n) # 2075 out of 2469 had only one screen
dat.resp.enc <- full_join(dat.resp.enc, screen.count, by="PAT_ID") %>% rename("n_encounter" = "n") # Append number of encounters to dataset

# Demographics of those who responded
response_ids <- dat.resp.enc %>% filter(sum > 0) %>% distinct(PAT_ID) %>% c() %>% .[[1]]
dat.resp <- dat %>% filter(PAT_ID %in% response_ids)
dat.resp.distinct <- dat.resp %>% distinct(PAT_ID, .keep_all = T)
```

```{r}
# Distribution of SDH screen counts
table(screen.count$n)
```

From this distribution of screening counts, a majority of individuals only had 1 screen, going up to 2 individuals having 9 screens.


\newpage
# Screening probability analysis

Next, we explored the chances of an individual receiving a screen, depending on their covariates. To distinguish individuals who have received a screen or not, I chose to use the FLO_MEAS_ID which indicates the type of SDH question being asked to the subject. If an individual does not receive a screen, this value is NA.

```{r}
# Individuals with more visits more likely to have at least 1 SDH screening
# Categories by visit: 0: 1, 1: 1-3, 2: 3-10, 3: 10-25, 4: 25-50, 5: >50 
round(tapply(!is.na(dat.distinct$FLO_MEAS_ID), list(dat.distinct$cat_visit), FUN = mean),2)

# Overall rate of having at least 1 SDH screening
mean(!is.na(dat.distinct$FLO_MEAS_ID))

# Older individuals less likely to have a screening
# Categories by age: 0: <32, 1: 32-44, 2: 44-58, 3: >58
round(tapply(!is.na(dat.distinct$FLO_MEAS_ID), list(dat.distinct$cat_age), FUN = mean),2)

# Probability of having at least 1 SDH screening by race and ethnicity
round(tapply(!is.na(dat.distinct$FLO_MEAS_ID), 
             list(dat.distinct$Race, dat.distinct$Ethnicity), FUN = mean),2)

# Probability of having at least 1 SDH screening by language
data.frame(probability=round(tapply(!is.na(dat.distinct$FLO_MEAS_ID), 
                  list(dat.distinct$Language), FUN = mean),2) ) %>% filter(probability>0)
```

\newpage
## Dates and screened demographics
We explored the range of dates for both last visit and encounter dates. Note that the ranges of these do not match up, indicating an issue with some encounter dates being missing. The NAs in this context also represent the individuals who never had an encounter/SDH screen. 

```{r}
# Range of last visit and encounter dates
summary(dat.distinct$Last_Visit_Date)
summary(dat.distinct$ENC_DATE)
```

We also analyzed various distributions of individuals who had at least 1 response:

```{r}
# Race by ethnicity table of those who responded
table(dat.resp.distinct$Race, dat.resp.distinct$Ethnicity)

# Age category table of those who responded: 0: <32, 1: 32-44, 2: 44-58, 3: >58
table(dat.resp.distinct$cat_age)

# Visit category table of those who responded: 0: 1, 1: 1-3, 2: 3-10, 3: 10-25, 4: 25-50, 5: >50 
table(dat.resp.distinct$cat_visit)
```

\newpage
# Response analysis

```{r, echo=FALSE}
getResponse <- function(questionType){
  dat.resp %>% filter(MEAS_DISP_NAME == questionType) %>% pull(RESPONSE) %>%
    table() %>% sort(decreasing = T) %>% as_tibble() %>% rename(!!questionType:=".")
}

questions <- dat$MEAS_DISP_NAME %>% unique() %>% .[-28] # All response types
lapply(questions,getResponse)
```










